---
title: "9041 Assignment 2"
author: Selena Esteves
date: today
format: 
    html:
      toc: true
      code-fold: true
      code-summary: "Show code"
      df-print: paged
execute:
    echo: false
    warning: false
---

```{r, load libraries}
#| output: FALSE
# Output false ensures your reproducible document does not show all of your code!

# Remember to install packages you do not already have

library(tidyverse)
library(readxl)
library(psych)
library(gt)
library(janitor)
```

```{r}
#| output: FALSE

# Load data
data<-read_excel("./bfi.xlsx")


# Inspect data for the big five dataset
head(data)
```

```{r, clean data}
#| output: FALSE

# Use mutate to rename variables in a meaningful way - such that those assigned gender value 1 in the dataset are listed as Male and those assigned gender value of 2 are listed as Female. Do the same for education values
# Extract for only variables age, Gender, education, and big five values
# Mutate to convert bfi values to integers

bfi.clean <- data |>
    mutate(Gender = case_when(gender==1 ~ "Male", gender==2 ~"Female", TRUE ~ "NA")) |> 
    mutate(education = case_when(education==1 ~ "High School", education==2 ~ "Finished High School",education==3 ~ "Some College", education==4 ~ "College Graduate", education==5 ~ "Graduate Degree", TRUE ~ "NA")) |> 
    select(age, Gender, education, A1:O5) |>
    mutate_at(4:28, as.integer)

```

```{r, score items}
#| output: FALSE

# Answer key: which items belong to each subscale.
# "-" marks reverse-coded items (e.g., E1 "Don't talk a lot" -> high = LESS extraverted), which get flipped so higher scores always mean more of the trait.

keys<- list(Openness=c("O1", "-O2", "O3", "O4", "-O5"),
                Conscientiousness=c("C1", "C2", "C3", "-C4", "-C5"),
                Extraversion=c("-E1", "-E2", "E3", "E4", "E5"),
                Agreeableness = c("-A1", "A2", "A3", "A4", "A5"),
                Neuroticism=c("N1", "N2", "N3", "N4", "N5"))


# Average items per person, and compute Cronbach's alpha.
# Returns $scores (per-person scores as averages since we set totals to F) and $alpha (reliability)
scores<-scoreItems(keys, bfi.clean, totals = F, min=1, max=6)

# Turn this into a dataframe!
scores.df<-as.data.frame(scores$scores)

```

```{r, combine data frames}
#| output: FALSE

# Combine the original data set with the new scores dataframe
bfi.clean<-cbind(bfi.clean, scores.df)

# Inspect new data set
head(bfi.clean)
```

```{r, reduced data}
#| output: FALSE

# Select only for the columns which you want to appear in your final data table
bfi.final<- bfi.clean |>
    select("age", "Gender", "education", "Openness", "Conscientiousness", "Extraversion", "Agreeableness", "Neuroticism") |>
    mutate_if(is.numeric, round, 2)
```

```{r, reproducible table for education distribution}
#| output: FALSE

# Create a demographics table of education by gender
demographictable <- bfi.final |>
  tabyl(education, Gender)

# Create your reproducible gt table
gttable.1 <- demographictable |>
  gt() |>
  tab_header(
    title = md("*Distribution of Males and Females by Education Level*")
  )

```

```{r, reproducible table for big five scores}
#| output: FALSE

# Round Cronbach's alpha values to 2 decimal places.
alphas <- t(round(scores$alpha, 2))

# Summary statistics for certain columns
summary <- describe(select(bfi.final, Openness:Neuroticism))
summary

# Label sample size for later in your reproducible document using describe
n <- (describe(bfi.final))[1,2]

# Name an object of Subscales so this can appear in your table
Subscales <- c("Openness", "Conscientiousness", "Extraversion", "Agreeableness", "Neuroticism")

# Create a column called Subscales in your summary table
summary$Subscales<- Subscales


# Select for only the columns which you want to appear in your final table
summary <- summary |>
    select(Subscales, mean, median, sd, range) |>
       mutate_if(is.numeric, round, 2)

# Combine summary information with alpha values
final.reptable <- cbind(summary, alphas)

# Create a reproducible table
gttable.2<- gt(final.reptable) |>
    tab_header(
        title = md("*Descriptive Statistics for the Big Five Inventory Subscales*")) |>
    tab_stubhead(label= "Composites")|>
    tab_spanner(
        label="Central Tendency",
        columns = c(mean, median)
    )|>
        tab_spanner(
        label="Variability",
        columns = c(sd, range)
        )|>
            tab_spanner(
        label="Reliability Values (Chronbach's alpha)",
        columns = c(alpha)
            ) 

```

```{r, filtered dataset}
#| output: FALSE

# Use filter to extract only those participants who are over 18 and under 65

bfi.filtered <- bfi.clean |>
    filter(age >= 18 & age <= 65)

# Checking that age restriction worked
table(bfi.filtered$age) 
```

```{r, reproducible FILTERED table for big five scores}
#| output: FALSE

# Round Cronbach's alpha values to 2 decimal places. t transposes the data, which is not needed with the other data set. Follow code in instruction document.
alphas_filtered <- t(round(scores$alpha, 2))

# Summary statistics for certain columns
summary_filtered <- describe(select(bfi.filtered, Openness:Neuroticism))
summary_filtered

# Name an object of Subscales so this can appear in your table
Subscales <- c("Openness", "Conscientiousness", "Extraversion", "Agreeableness", "Neuroticism")

#Create a column called Subscales in your summary table
summary_filtered$Subscales<- Subscales

# Select for only the columns which you want to appear in your final table
summary_filtered <- summary_filtered |>
    select(Subscales, mean, median, sd, range) |>
       mutate_if(is.numeric, round, 2)

# Combine summary information with alpha values
filtered.reptable <- cbind(summary_filtered, alphas_filtered)

# Create a reproducible table
gttable.3<- gt(filtered.reptable) |>
    tab_header(
        title = md("*Descriptive Statistics for the Big Five Inventory Subscales in people 18 to 65*")) |>
    tab_stubhead(label= "Composites")|>
    tab_spanner(
        label="Central Tendency",
        columns = c(mean, median)
    )|>
        tab_spanner(
        label="Variability",
        columns = c(sd, range)
        )|>
            tab_spanner(
        label="Reliability Values (Chronbach's alpha)",
        columns = c(alpha)
            ) 
```
## Intro

## Methods

### Participants

In this data set there were `{r} n` participants, with varying distributions of education type (Table 1).

`{r} gttable.1`

### Measures

In this BFI data set, the 5 factors refer to the 5 composites included in the Big Five Inventory. Openness, which indicates one's open-mindedness, creativity, and likelihood to embrace new experiences (Cronbach's alpha = `{r} alphas[1]`). Conscientiousness, which refers to one's tendency to be reliable, organized, and disciplined (Cronbach's alpha = `{r} alphas[2]`). Extraversion, which refers to one's social interactions and emotional expression (Cronbach's alpha = `{r} alphas[3]`). Agreeableness, which refers to one's tendency to cooperate or desire to compete (Cronbach's alpha = `{r} alphas[4]`). and Neuroticism which refers to one's tendency toward depression, anxiety, or negative feelings (Cronbach's alpha =`{r} alphas[5]`).

### Data Analysis Strategy

The data for variables O2, C4, E1, E2, and A1 were reverse keyed.

## Results

`{r} gttable.2`
In this BFI data set, the 5 factors refer to the 5 composites included in the Big Five Inventory. Participants had a mean score on Openness of `{r} final.reptable[1,2]`, on Conscientiousness of `{r} final.reptable[2,2]`, on Extraversion of `{r} final.reptable[3,2]`, on Agreeableness of `{r} final.reptable[4,2]`, and on Neuroticism of `{r} final.reptable[5,2]`.

`{r} gttable.3`
When only including participants over age 18 and below age 65, this BFI data set, the 5 factors refer to the 5 composites included in the Big Five Inventory. Participants had a mean score on Openness of `{r} filtered.reptable[1,2]`, on Conscientiousness of `{r} filtered.reptable[2,2]`, on Extraversion of `{r} filtered.reptable[3,2]`, on Agreeableness of `{r} filtered.reptable[4,2]`, and on Neuroticism of `{r} filtered.reptable[5,2]`.

## Discussion
